Ethnicity sensitive author disambiguation using semi-supervised learning

### 2016-04-29 paper/author_disambiguation.tex
rephrase to keep 15 pages as before

In this work, we have revisited and validated the general author disambiguation
pipeline introduced in previous independent research work.
The generic approach is composed of three components, whose design and tuning are all critical
to good performance: (i) a blocking function for pre-clustering signatures
and reducing computational complexity, (ii) a linkage function for identifying
signatures with coreferring authors and (iii) the agglomerative clustering of
signatures. Making use of a distinctively large dataset of more than 1 million
crowdsourced annotations, we experimentally study all three components and
propose further improvements. With regards to blocking, we suggest to use
phonetization of author names to increase recall while maintaining low
computational complexity. For the linkage function, we introduce
ethnicity-sensitive features for the automatic tailoring of disambiguation to non-Western
author names whenever necessary. Finally, we explore semi-supervised cut-off
threshold strategies for agglomerative clustering. For all three components,
experiments show that our refinements all yield significantly better author
disambiguation accuracy.

Overall, these results all encourage further improvements and research. For
blocking, one of the challenges is to manage signatures with inconsistent
surnames or first given names (cases 4 and 5, as described in
Section~\ref{methods:blocking}) while maintaining blocks to a tractable size.
As phonetic algorithms are not yet perfect, another direction  for further work is the design of better
phonetization functions, tailored for author disambiguation. For the linkage function,
the good results of the proposed features pave the way for further research  in
ethnicity-sensitive author disambiguation. The automatic fitting of the
pipeline to cultures and ethnic groups for which standard author disambiguation
is known to be less efficient (\eg, Chinese authors with many homonyms)  indeed
constitutes a direction of research with great potential benefits for the
concerned scientific communities.

Moreover, the techniques presented in this work can be easily adapted to the broader
problem of named entity disambiguation and might improve accuracy of semantic search
algorithms.

>>>>

In this work, we have revisited and validated the general author disambiguation
pipeline introduced in previous independent research work.
The generic approach is composed of three components, whose design and tuning are all critical
to good performance: (i) a blocking function for pre-clustering signatures
and reducing computational complexity, (ii) a linkage function for identifying
signatures with coreferring authors and (iii) the agglomerative clustering of
signatures. Making use of a distinctively large dataset of more than 1 million
crowdsourced annotations, we experimentally study all three components and
propose further improvements. With regards to blocking, we suggest to use
phonetization of author names to increase recall while maintaining low
computational complexity. For the linkage function, we introduce
ethnicity-sensitive features for the automatic tailoring of disambiguation to non-Western
author names whenever necessary. Finally, we explore semi-supervised cut-off
threshold strategies for agglomerative clustering. For all three components,
experiments show that our refinements all yield significantly better author
disambiguation accuracy.
In general, the results encourage further improvements and research. For blocking, one of the challenges is to manage signatures with inconsistent surnames or first given names (cases 4 and 5, as described in Section~\ref{methods:blocking}) while maintaining blocks to a tractable size. As phonetic algorithms are not yet perfect, another direction  for further work is the design of better phonetization functions, tailored for author disambiguation. For the linkage function, the good results of the proposed features pave the way for further research in ethnicity-sensitive author disambiguation. The automatic fitting of the pipeline to ethnic groups is better that a standard one that is less efficient for author names with many homonyms (\eg,Chinese authors). Collecting a bigger name-to-ethnicity dataset with better coverage of names is another future work to consider.
Moreover, the techniques presented in this work can be easily adapted to the broader
problem of named entity disambiguation and might improve accuracy of semantic search
algorithms.
